---
title: "GuidedMod11"
author: "Sarah Hall"
date: "2024-11-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(faraway)
Data<-wcgs
set.seed(6021) ##for reproducibility to get the same split
sample<-sample.int(nrow(Data), floor(.50*nrow(Data)), replace = F)
train<-Data[sample, ] ##training data frame
test<-Data[-sample, ] ##test data frame

```


Before fitting a model, create some data visualizations to explore the relationship
between these predictors and whether a middle-aged male develops coronary heart
disease.

```{r}
chart1<-ggplot2::ggplot(train, aes(x=age, fill=chd))+
  geom_bar(position = "fill")+
  labs(x="Age", y="Proportion",
       title="Proportion of CHD by Age")

chart2<-ggplot2::ggplot(train, aes(x=sdp, fill=chd))+
  geom_bar(position = "fill")+
  labs(x="SDP?", y="Proportion",
       title="Proportion of CHD by SDP")

chart3<-ggplot2::ggplot(train, aes(x=dbp, fill=chd))+
  geom_bar(position = "fill")+
  labs(x="DBP?", y="Proportion",
       title="Proportion of CHD by DBP")

chart4<-ggplot2::ggplot(train, aes(x=cigs, fill=chd))+
  geom_bar(position = "fill")+
  labs(x="Cigarettes", y="Proportion",
       title="Proportion of CHD by Cigs")

chart5<-ggplot2::ggplot(train, aes(x=dibep, fill=chd))+
  geom_bar(position = "fill")+
  labs(x="Dibep", y="Proportion",
       title="Proportion of CHD by Dibep")

library(gridExtra)
gridExtra::grid.arrange(chart1, chart2, chart3, chart4, chart5, ncol = 2, nrow = 3)

```

```{r}
#Two way tables of dibep and DBP
prop.table(table(train$chd,train$dibep),1)
prop.table(table(train$chd,train$dbp),1)
```

2. Use R to fit the logistic regression model using all the predictors listed above, and
write the estimated logistic regression equation.

```{r}
result<-glm(chd~age+sdp+dbp+cigs+dibep, family=binomial,  data=train)
summary(result)
```
y=0.06x1 + 0.02x2 + 0.01x3+0.02x4-0.53x5 - 8.309

1
3. Interpret the estimated coefficient for cigs in context.

The number of cigarettes smoked per day are significant and have a positive association with chd.

4. Interpret the estimated coefficient for dibep in context.

Dibep is significant and have a positive association with chd.

5. What are the estimated odds of developing heart disease for an adult male who is
45 years old, has a systolic blood pressure of 110 mm Hg, diastolic blood pressure
of 70 mm Hg, does not smoke, and has type B personality? What is this person’s
corresponding probability of developing heart disease?

```{r}
(45*.06) + (110*.015) + (70*.012) + .021 -.53 -8.31

```



```{r}
library(faraway)
faraway::vif(result)
```



6. Carry out the relevant hypothesis test to check if this logistic regression model with
five predictors is useful in estimating the odds of heart disease. Clearly state the null
and alternative hypotheses, test statistic, and conclusion in context.

7. Suppose a co-worker of yours suggests fitting a logistic regression model without the
two blood pressure variables. Carry out the relevant hypothesis test to check if this
model without the blood pressure variables should be chosen over the previous model
with all four predictors.

```{r}
reduced<-glm(chd~age+cigs+dibep, family=binomial, data=train)

##test to compare reduced and full model
##test stat
TS<-reduced$deviance-result$deviance
TS
```
```{r}
##pvalue
1-pchisq(TS,2)
```
```{r}
qchisq(1-0.05,4)
```
We fail to reject the null hypothesis



8. Based on the Wald test, is diastolic blood pressure a significant predictor of heart
disease, when the other predictors are already in the model?




9. Based on all the analysis performed, which of these predictors would you use in your
logistic regression model?